---
title: "Final Project"
author: "Hannah Bolton, Vidhya Narayanan, Aishi Sharma, Blake Needham"
date: "2023-12-01"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(data.table)
library(R.utils)
library(tidyverse)
library(ggplot2)
library(extrafont)
library(knitr)
```

# 1. Introduction  
Attention deficit hyperactive disorder (ADHD) stands as a prevalent mental
disorder among children, characterized by symptoms like difficulty focusing,
excessive movement, and impulsive behavior. Its chronic nature significantly
affects various facets of life such as academic performance, professional
success, relationships, and daily functioning. When left unaddressed, ADHD can
detrimentally impact a child's self-esteem and social abilities. Roughly 8.4% of
children and 2.5% of adults are estimated to have ADHD, a condition often
recognized in school-aged kids due to classroom disruptions or academic
struggles. Diagnosis rates are higher among boys due to variations in symptom
expression, with boys more prone to hyperactivity and outward symptoms, while
girls might display inattentiveness. However, this doesn't imply a higher
likelihood of ADHD in boys; it reflects differences in how symptoms manifest.

The way that ADHD is portrayed in the media frequently leads to the idea that it
is an impulsive, hyperactive disorder. They could find it difficult to remain
motionless, disrupt others, and act impulsively without thinking through the
repercussions. ADHD is also characterized by inattentiveness. People suffering
from ADHD may find it difficult to focus, forget things readily, and misplace
things easily. Many representations emphasize the difficulties that people with
ADHD have in the classroom. Their inability to concentrate, stay organized, and
finish assignments could cause them to suffer academically. Positively speaking,
some representations highlight the inventiveness and creativity connected to
ADHD. Some people with ADHD might be excellent artists or do well in fast-paced
environments. Gender variations in ADHD presentation may occasionally be
reflected in representations. Although girls with ADHD may present with more
inattentive symptoms than boys, this can sometimes result in an underdiagnosis,
or an incorrect diagnosis, which is very dangerous for everyone involved. Stigma
surrounding the illness leads to some people hiding thier diagnosis out of fear
of being treated differently from their peers.

#### *Relevance to each team member*
Blake -  
I have a brother who is diagnosed with ADHD as well as Dyslexia. He handles both
of these together with support from our family. His diagnoses came recently, so
we are all getting adjusted to this new way of life, and figuring out how we can
all best support him.  
Aishi -  
I grew up around a lot of individuals who were diagnosed with ADHD as a child. I
wanted to explore the topic of the disparity of ADHD diagnoses after discussing
the topic of disparity of ADHD diagnoses with Hannah.  
Hannah -  
I was diagnosed with ADHD in March which is why this topic is highly prevalent
in my life. I also recently wrote a paper discussing the disparities in ADHD
diagnoses among people of different races.  
Vidhya -  
When I was writing my “better information world” essay for Info 200, I wound up
doing research on how different racial minorities in particular were negatively
impacted by a lack of resources to diagnose different disorders. When I was
given this project, it reminded me of that scenario, so after communicating with
my group we realized that ADHD is a very relevant topic to conduct research on.

Even though ADHD is prevalent among developmental disorders, it remains widely
misunderstood, leading to potential harm for those affected. Misconceptions
surrounding ADHD, like doubting its validity as a treatable condition or
assuming unnecessary medication usage in children, can discourage individuals
from seeking necessary treatment. A recent study in 2021 highlighted the
pervasive impact of these myths on the self-perception of those with ADHD,
contributing to reluctance to seek help or fostering feelings of guilt and
shame. Shockingly, a 2019 study revealed that individuals with ADHD are
approximately three times more prone to suicide attempts than the general
population. When these stigmas and misconceptions impede access to proper
treatment, the consequences can be tragically life-threatening.

## Data

The datasets we are using come directly from the Center for Disease Control and
Prevention (CDC), which is a well-known and reputable source for medical data.
We decided to use these datasets because the website allowed us to customize
what years and categories of data we were interested in, so out of different
mental disorders we were able to focus on ADHD specifically, which was not an
option on the other websites that we investigated. The ability to pick the year
also proved to be very convenient, because one of the essential elements of our
investigation was the ability to compare parallel data between 2000 and 2020.
The website allowed us to download two CSV files, one from 2000, and one from
2020. The datasets originate from a survey done using data from a variety of
households across the country, collecting data about each individual living
there. The Data collection was done by the Center for Disease Control (CDC) is
this federal organization is a division of the Department of Health and Human
Services in the United States. The main goal of the CDC, which has its main
office in Atlanta, Georgia, is to safeguard public health and safety by
preventing and controlling illnesses, accidents, and disabilities. This data can
be publicly accessed by anyone with adequate internet access. There are not any
explicit restrictions but the website could prove to be difficult for someone
with minimal experience navigating databases and interpreting complex data
structures.

The datasets contain information on 12,514 people, if they have an ADHD
diagnosis, their race, and their household incomes, with samples coming from
both the years 2000 and 2020.

We only ran into a few issues with data quality in our work. There were very few
data points for certain races such as Native and African Americans. This became
apparent in a few of the calculations that we made. For example, in one of the
plots we made, we intended to find what percentage of each race, depending on
income group had ADHD. The data stated that 50% of Native Americans above the
income level of $100,000 had ADHD. This data seemed to be a potential outlier,
so when we investigated what the issue could be, we noticed that only 39 Native
Americans were interviewed. This is an incredibly low number compared to the
6,868 white people who were surveyed. In addition to the inconsistent
proportions of each race represented in the data, the majority of the data
points in the original datasets had not been asked the question regarding ADHD
diagnosis, so we had to filter them out to provide accurate and relevant data.
Additionally, in approximately 40% of the data points in the original dataset,
there were five race values regarding “unknown race.” These pieces of data
proved to be irrelevant to our data analysis because our report is contingent on
the racial identities of the people we are observing, so we excluded them.

Each row in our final dataset represents the data about one racial group in a
certain year that belongs to a certain income group, and whether or not they
have ADHD. We have 12,514 observations in our merged dataset of the 2000 and
2020 data.

The relevant variables to our dataset are the lower bound of household income
levels, the upper bound of household income levels, racial category, whether
they are white or non-white, gender, year, and the type of ADHD diagnosis. The
year of collection is called ‘year’ and is coded with numerical values. The
gender is called ‘sex’ and is coded by either ‘M’ or ‘F’. The racial category is
called ‘race’ and is coded by several different values: ‘White’, ‘Black’,
‘Native’, and ‘Asian’. The white vs. non-white variable is called ‘whiteOrNot’,
and is coded by either ‘white’ or ‘non-white’. The variable for the type of ADHD
diagnosis is called ‘diagnosisStatus’ and is coded by three different values:
‘No-ADHD’, ‘Has-ADHD’, and ‘Untested’. The variable for the lower bound of
household income level is called ‘lowerBoundIncome’ and the variable for the
upper bound of household income level is called ‘upperBoundIncome’; both are
coded by numerical values representing USD.

We did not have any missing, non-plausible, or other quality issues in our final
dataset, other than a low number of certain races interviewed at all, in
particular Native Americans.

```{r}
# loading the datasets
data2000 <- read.delim("data/adhd race income 2000.csv", sep = ",")
data2020 <- read.delim("data/adhd race income 2020.csv", sep = ",")

# cleaning up our 2000 dataset
filter2000 <- data2000 %>%
  select(YEAR, RACENEW, SEX, INCFAM97ON2, ADDEV) %>%
  rename(year = YEAR, race = RACENEW, sex = SEX)
filter2000 <- filter2000 %>%
  mutate(
    lowerBoundIncome = case_when(
      INCFAM97ON2 == 10 ~ 0,
      INCFAM97ON2 == 20 ~ 35000,
      INCFAM97ON2 %in% c(30, 31) ~ 75000,
      INCFAM97ON2 == 32 ~ 100000
    ),
    upperBoundIncome = case_when(
      INCFAM97ON2 == 10 ~ 34999,
      INCFAM97ON2 == 20 ~ 74999,
      INCFAM97ON2 %in% c(30, 31) ~ 99999,
      INCFAM97ON2 == 32 ~ 100000
    )
  ) %>%
  filter(INCFAM97ON2 %in% c(10, 20, 30, 31, 32)) %>% 
  select(-INCFAM97ON2)  
filter2000 <- filter2000 %>%
  filter(race %in% c(100, 200, 300, 400)) %>%
  mutate(
    race = case_when(
      race == 100 ~ "White",
      race == 200 ~ "Black",
      race == 300 ~ "Native",
      race == 400 ~ "Asian"
    ),
    whiteOrNot = ifelse(race == "White", "white", "non-white")
  )
filter2000 <- filter2000 %>%
  filter(sex %in% c(1, 2)) %>%
  mutate(sex = ifelse(sex == "1", "M", "F"))
filter2000 <- filter2000 %>%
  filter(ADDEV %in% c(1, 2, 9)) %>%
  mutate(
    diagnosisStatus = case_when(
      ADDEV == 1 ~ "No-ADHD",
      ADDEV == 2 ~ "Has-ADHD",
      ADDEV == 9 ~ "Untested"
    )
  ) %>%
  select(-ADDEV)

# cleaning up the 2020 dataset
filter2020 <- data2020 %>%
  select(YEAR, RACENEW, SEX, INCFAM07ON, ADDEV) %>%
  rename(year = YEAR, race = RACENEW, sex = SEX) %>%
  mutate(
    lowerBoundIncome = case_when(
      INCFAM07ON == 11 ~ 0,
      INCFAM07ON == 23 ~ 75000,
      INCFAM07ON %in% c(12, 20) ~ 35000,
      INCFAM07ON == 24 ~ 100000
    ),
    upperBoundIncome = case_when(
      INCFAM07ON == 11 ~ 34999,
      INCFAM07ON == 23 ~ 100000,
      INCFAM07ON %in% c(12, 20) ~ 74999,
      INCFAM07ON == 24 ~ 1000000
    )
  ) %>%
  filter(INCFAM07ON %in% c(11, 12, 20, 23, 24)) %>%  
  select(-INCFAM07ON) 
filter2020 <- filter2020 %>%
  filter(race %in% c(100, 200, 300, 400)) %>%
  mutate(
    race = case_when(
      race == 100 ~ "White",
      race == 200 ~ "Black",
      race == 300 ~ "Native",
      race == 400 ~ "Asian"
    ),
    whiteOrNot = ifelse(race == "White", "white", "non-white")
  )
filter2020 <- filter2020 %>%
  filter(sex %in% c(1, 2)) %>%
  mutate(sex = ifelse(sex == "1", "M", "F"))
filter2020 <- filter2020 %>%
  filter(ADDEV %in% c(1, 2, 9)) %>%
  mutate(
    diagnosisStatus = case_when(
      ADDEV == 1 ~ "No-ADHD",
      ADDEV == 2 ~ "Has-ADHD",
      ADDEV == 9 ~ "Untested"
    )
  ) %>%
  select(-ADDEV)

# merging the datasets so we are able to compare the two years
mergedData <- rbind(filter2000, filter2020)
write.csv(mergedData, "Data/Final Cleaned Data.csv", row.names=FALSE)
```

Here is a sample of our dataset:
```{r sample dataset}
kable(mergedData[sample(nrow(mergedData), 10),])
```

# 2. Methods

From the original datasets, we included the variables ‘YEAR’, ‘RACENEW’, ‘SEX’,
‘ADDEV’, and either ‘INCFAM97ON2’ or ‘INCFAM07ON’, depending on the dataset. For
ease and continuity of reading and writing variable names, we changed some of
the variable names; ‘YEAR’ became ‘year’, ‘RACENEW’ to race, and ‘SEX’ to ‘sex’.
We did not change the names of ‘ADDEV’, ‘INCFAM97ON2’, and ‘INCFAM07ON’ because
we knew we were going to modify them further and would be able to replace those
values at a later point in the process.

Once selecting the variables we knew we wanted to work with, we needed to modify
a few variables. The income variables, ‘INCFAM97ON2’ and ‘INCFAM07ON’, belonging
to the 2000 and 2020 datasets respectively, had generic value placeholders that
represented different income ranges, and the ranges differed between years
because they altered their question after 2007. We decided to modify these
representations of household income by creating two new variables:
‘lowerBoundIncome’ and ‘upperBoundIncome’. These variables are coded as
numerical values representing the lower and upper bounds of household income
that were specified in the original datasets’ codebooks. When doing this, we had
to create a couple generic values: for any ‘unspecified’ options, we represented
those by the number -999, and for any income range without an upper bound we
represented the upper bound with 1,000,000 as that number is far higher than any
of the other specified values.

The variable ‘ADDEV’ was also coded with generic numerical placeholders
representing different diagnosis statuses. As mentioned previously, a majority
of the observations in both datasets had not even been asked the question
regarding ADHD, so we had to filter out any observations with that value in
‘ADDEV’ to be able to comprehensively analyze the data. We then converted the
numerical placeholders to the more easily comprehendible values of ‘No-ADHD’,
‘Has-ADHD’, and ‘Untested’, and renamed the variable ‘diagnosisStatus’.

We also wanted to make the ‘race’ column more easily understandable, as it too
was coded as numerical placeholders for different racial identifications. Also
as mentioned previously, there were several values that were some version of
“race unknown”, so we decided to cut out any observations with those values in
the race column as one of the main points of analysis in our project is race.
We acknowledged that this could potentially be exclusionary, but it is difficult
to analyze data with many unknown values, particularly when those values are for
one of the main areas of examination. We then converted the numerical
placeholders to ‘White’ for white-identifying individuals, ‘Black’ for Black- or
African American-identifying individuals, ‘Native’ for Native
American-identifying individuals, and ‘Asian’ for Asian-identifying individuals.

We combined the data for our 2000 and 2020 results into a data frame called
‘mergedData’. We then created a summarization data frame called
‘adhdPercentByRacePop’ where we grouped our data by year, race, and diagnosis
status and then counted the total population of each race, as well as the
population of each race that fell into each diagnosis category. We included six
columns: year, race, diagnosisStatus, the number of people in that diagnosis
status column (count), the RacePopulation, and then the percentage of count to
RacePopulation (relativeToRacePopulation). This let us understand the prevalence
of ADHD in each racial group based on the relative population for each racial
group, as well seeing which groups were over/underrepresented in data collection.

We selected our sample by using the data source that seemed most trustworthy,
which was the CDC. Then, we decided on the years 2000 and 2020 because we felt
it was a big enough time difference to examine any changes over time and those
years had a lot of research conducted.

A majority of the observations in both datasets had not even been asked the
question regarding ADHD, so we had to filter out any observations with that
value in ‘ADDEV’ to be able to comprehensively analyze the data. We then
converted the numerical placeholders to the more easily comprehendible values of
‘No-ADHD’, ‘Has-ADHD’, and ‘Untested’, and renamed the variable
‘diagnosisStatus’. We also removed many of the “unsure about race” values
because those would have just taken up space in our data tables.

By removing a lot of the data, especially for the race column, those groups who
were unsure lost their representation from the original data-set to ours, and
maybe those groups would’ve helped us uncover significant statistics about the
relationship between race, income, levels and ADHD diagnosis.

# 3. Results

We started with a dataset that included many columns besides the ones we chose,
including columns for height and weight. Although these variables could
contribute to an interesting discussion, our research is primarily focused on
the relationship between race, ADHD diagnosis, and income level, so those
variables were irrelevant to us. After getting rid of those columns, we changed
the data values inside each of the columns. The values at first needed to be
deciphered using a key, so we adjusted the data to match the values in the key
so that the data frame can be interpreted by the casual reader. We also needed
to adjust the data frame to match the manipulation requirements. We needed a new
categorical variable, so we created a column called “whiteOrNot” that classified
the races of each observation as either “white” or “non-white”. We needed a new
numerical variable, so we took the character value that defined each income
level and replaced it with two numerical columns, one representing the lower
bound value of the income group and one representing the upper bound value. For
our summarization data frame, we created a data frame that contained the number
of people in each race + income group that had each type of adhd grouping (has
adhd, doesnt have it, or unknown) in a column named count. For reference, the
data frame also included the total population of observations in each race so
readers can understand what the number of people in each adhd grouping compared
to the total population of each group would look like proportionally in a column
called racePopulation. In the final column titled relativeToRacePopulation, we
did the operation count/racePopulation*100 to give us the percentage of people
in each race & income croup that belong to each adhd grouping out of the total
population.

```{r population percentages, message = FALSE, warning = FALSE}
# creating a dataframe that shows the percentage of diagnosed/undiagnosed/unknown
# individuals per the total number of samples for that race
adhdPercentByRacePop <- mergedData %>%
  group_by(year, race, diagnosisStatus) %>%
  summarize(
    count = n(),
  ) %>%
  bind_rows( 
    mergedData %>%
      filter(whiteOrNot %in% c("Black", "Native", "Asian")) %>%
      group_by(year, race, diagnosisStatus) %>%
      summarize(
        count = n(),  
      )
  ) %>%
  arrange(desc(race)) %>%
  ungroup() %>% 
  group_by(year, race) %>%  
  mutate(
    racePopulation = sum(count),
    relativeToRacePopulation = count / racePopulation * 100
  )

# creating a dataframe that shows the percentage of diagnosed/undiagnosed/unknown
# individuals per the total number of samples
adhdPercentByTotalPop <- mergedData %>%
  group_by(year, race, diagnosisStatus) %>%
  summarize(
    count = n(), 
  ) %>%
  bind_rows(
    mergedData %>%
      filter(whiteOrNot %in% c("Black", "Native", "Asian")) %>%
      group_by(year, race, diagnosisStatus) %>%
      summarize(
        count = n(), 
      )
  ) %>%
  arrange(desc(race)) %>%
  ungroup() %>%  
  group_by(year) %>%  # 
  mutate(
    TotalPopulation = sum(count),
    popPercent = count / TotalPopulation * 100
  )
```

First, we created a bar graph that graphed the percentage of people out of the
population of each race that had ADHD vs no ADHD vs. unknown for the years 2000
and 2020. There are four different graphs for each race and two bar-columns on
each one representing each year.

```{r}
ggplot(mergedData, aes(x = factor(year), fill = diagnosisStatus)) +
  geom_bar(position = "fill") +
  facet_wrap(~race) +
  labs(title = "Percentage of each race's population with each ADHD status, 2000 vs. 2020",
       x = "Year",
       y = "Percentage of People",
       fill = "ADHD Status")  
```

In this graph, it is apparent that the presence of ADHD in 2000 was highest for
the white population, but in 2020 there was a drastic spike in ADHD rates for
Native individuals, making up about 25% of their total population. However, when
we cross-referenced this with our summary data frame, we noticed that in 2020,
Native people only had 62 values while White people had about 7,000, meaning
that the data would probably not be as accurate. However, we did notice a bit
of an increase in the ADHD levels of the Black population as well, matching the
percentage of white people with an ADHD diagnosis in 2020.

We wanted to see if there was any relationship between this and income levels,
so we created a second graph that measured the percentage of people in each race
that belonged to each income group each year.

```{r percent of people in each race by income group}
mergedData$lowerBoundIncome <- factor(mergedData$lowerBoundIncome, 
                                      levels = c("0", "35000", "75000", "1e+05"),
                                      labels = c("0-35,000", "35,000-75,000", "75,000-100,000", "100,000+"))

# Create a bar plot with updated income group names
ggplot(mergedData, aes(x = factor(year), fill = factor(lowerBoundIncome))) +
  geom_bar(position = "fill") +
  facet_wrap(~race) +
  labs(title = "Percentage of each race that belongs to each income group, 2000 vs. 2020",
       x = "Year",
       y = "Percentage of People",
       fill = "Income Group") 
```

Although we concluded that our data for Native Americans was slightly flawed,
since we are only comparing it in reference with data from the same set of
people, we figured comparing the data about Native Americans from the first
graph to the second would not be detrimental to the accuracy of the patterns we
find. For both Black people and Native Americans, as the rate of people in the
0-35000 income category and 35000-75000 stayed relatively high, there was also a
higher level of ADHD diagnosis. Although Asian Americans and white people had
nearly identical income patterns in 2000 & 2020, the diagnosis rates were
significantly lower for Asian Americans, and there were also a lot more white
people used as data points than Asian Americans.

Finally, we made a graph that combined both of our previous graphs: It compared
the percentages of people in each income level per race with ADHD each year.
This means that there were eight graphs total, with four for each race in 2000
and then 2020.

```{r percentage of people with adhd in each income group per race, fig.width=16, fig.height=10}
merge2000 <- mergedData %>% filter(year == 2000)
percentADHD2000 <- ggplot(merge2000, aes(x = factor(lowerBoundIncome), fill = diagnosisStatus)) +
  geom_bar(position = "fill") +
  facet_wrap(~race) +
  labs(title = "Percentage of each race, depending on income group, that belongs to each ADHD diagnosis group in 2000",
       x = "Income Group",
       y = "Percentage of People",
       fill = "Diagnosis Status") +
  scale_fill_manual(values = c("No-ADHD" = "salmon", "Has-ADHD" = "darkblue", "Untested" = "lightgreen")) 

merge2020 <- mergedData %>% filter(year == 2020)
percentADHD2020 <- ggplot(merge2020, aes(x = factor(lowerBoundIncome), fill = diagnosisStatus)) +
  geom_bar(position = "fill") +
  facet_wrap(~race) +
  labs(title = "Percentage of each race, depending on income group, that belongs to each ADHD diagnosis group in 2020",
       x = "Income Group",
       y = "Percentage of People",
       fill = "Diagnosis Status") +
  scale_fill_manual(values = c("No-ADHD" = "salmon", "Has-ADHD" = "darkblue", "Untested" = "lightgreen"))

print(percentADHD2000)
print(percentADHD2020)
```

In these graphs, we realized that, at least with the data given, there was not a
very strong correlation between the income level and ADHD rate that could be
applied across all groups. For Black people, there was a decently strong trend
that grew in 2020 showing that lower income levels lead to higher rates of ADHD
diagnosis, but on the other hand, Native Americans also had similar income
levels and they saw the highest levels of ADHD in the 100,000+ population (but
we must remember the data for the Native Americans has certain limitations). For
Asian & White people, there was only a jump in ADHD levels primarily for lower
income groups in the year 2020, but in 2000 they were about equal. However, one
thing that was true was that ADHD diagnosis rates went up for *all* races from
2000 to 2020.

# 4. Findings

### Our key findings:
* There is a slight relationship between groups that have a higher number of
“lower” income individuals and a higher rate of ADHD  diagnosis.
    + Although race and income are possible contributing factors, it is important to
note that a higher rate of diagnostic checks for ADHD can lead to a higher
percentage of the population being screened as having ADHD.

* It is not feasible to determine one relationship between ADHD diagnosis rates
and income that fits all races for 2000 and 2020. However, it is safe to say
that as the rates of ADHD grew from 2000 to 2020, lower income groups were more
likely to also have a larger increase in ADHD for most racial groups excluding
Native Americans.

One limitation we faced was a lack of feasible data points for the Native
American racial group despite having sufficient data for the other races. I
believe our claim is decently substantiated and could be extended to serve as a
starting point/catalyst for future research in the field, even if not good
enough to be concrete evidence right now on a higher-tier research level.

For future work, it would be ideal to have a dataset that has a semi-equal
number of observations from each racial group and income level. This would, to a
high extent, get rid of any doubt about a certain group of individuals having
more representation / better data than another one. It would also be helpful to
use more complex algorithms made specifically for high level data science, and
not just the simple percentage based graphs we created using our level of R
knowledge.

# 5. Summary and Conclusion

This project was very meaningful in gaining practice in analyzing datasets and
figuring out ways to conclude, as well as evaluating the accuracy of the data
and those conclusions. In cleaning data, we learned to negotiate with each other
on one data would be significant and necessary to our investigation, and this
allowed us to grab some meaningful intake.

Through our summarization data frame, we were able to pinpoint some of the
shortcomings of our data, such as the fact that the White population was grossly
overrepresented, where as the native sample size was about less than 1% of
that. Although we did not completely discard our data for Native Americans
because of this fact, we did make sure to be conscious of the fact that the data
may not be as reliable as with the other races.

When we first compared the percentages of ADHD diagnosis for each race is 2000
vs. 2020, we noticed that the percentage was highest for white people in 2000,
but in 2020 there was a spike in the Native American ADHD rate and the rate for
Black People. So, to see if there was any underlying economic cause for this, we
also created a graph that compared the percentages of each race that eblonged to
each income group for both years. We noticed that generally, the Black and 
Native Americans tended to experience an augmentation/maintenance of the
percentage of people in the “lower” income categories, whereas for White and
Asian people the trend went down. Finally, to compare both our racial
percentages and income percentages, we created a bar graph that investigated the
rate of ADHD diagnosis for each race per income group. In this, we noticed that
for Even though Asians and White folk had nearly identical income statistics,
white people had higher levels of diagnosis, but those two groups shared the
quality of staying low and not having a spike in growth. However, for the Black
and Native lower income group there was a notable increase in ADHD rates from
2000 to 2020 which ended up overtaking the white percentage. For Asian Americans
as well there was a slightly higher rate of ADHD for the 0-35,000 income
category, but it was still quite minimal.

We were able to conclude that **although there is not one perfect trend that
fits all races for a relationship between income and ADHD diagnosis rates, as
the rates of ADHD grew from 2000 to 2020, lower income groups were more likely
to also have a larger increase in ADHD**. This study opened our eyes to the fact
that not all racial and income groups have equal access to ADHD information and
screening within their communities, one of the side effects being that we were
not able to conduct as precise of an investigation as we wanted. The results of
this investigation could be used by school districts and medical centers, as
they could figure out which geographical areas they need to put their focus into
for understanding the causes of ADHD and screening based on income. Using the
slight patterns we found between ADHD diagnosis and income for certain races,
these organizations could also conduct a deeper dive into that, and see if there
are any environmental issues that could be provoking this. These results could
also be used by other researchers to inspire a future research paper on the
topic.